Spatial variation of the native colon microbiota in healthy adults

### 2018-01-30 submission/manuscript.Rmd
paper discussion revisions

We detected _F. nucleatum_ and _P. asacharolytica_ in 8 and 5 of our subjects, respectively. These bacteria have been shown to be predictive of colorectal cancer in humans @Baxter2016 and have oncogenic properties in cell culture and in mice @Sears2014. Interestingly, while _F. nucleatum_ was found on both sides of the colon, _P. asacharolytica_ was only detected in the distal mucosa. Not much is known about the distribution of _P. asacharolytica_ along the colon, but given its anaerobic lifestyle and asacharolytic metabolism, it is not surprising that it resides in the less-oygen-rich and protein-rich distal mucosa @Albenberg2014. In studies examining bacteria on colorectal cancer tumors, _F. nucleatum_ was more commonly detected on proximal-sided tumors, and distribution of _F. nucleatum_ decreased along the colon to rectum @Mima2016. Of the 8 (40%) individuals positive for _F. nucleatum_ in the present study, the bacterium was spread across the proximal mucosa, distal lumen and distal mucosa. Data examining bacterial biofilms on the mucosa of CRC tumors suggests that _Fusobacteria_ species are more commonly found on proximal tumors and in biofilms, indicating that it is not only the presence of the bacterium but the structure of the tumor community that contributes to _Fusobacterium's_ role in tumorigenesis @Dejea2014. Finally, _Fusobacterium_ and _Porphyromonas_ populations not only co-occur on CRC tumors but also to synergize to promote tumorigenesis in an oral cancer model [@Whitmore2014; @Flynn2016]. Further analysis of the distribution and activities of these pathogens along the colon is needed to elucidate a mechanism for development of CRC or IBD subtypes in the proximal or distal colon.

The _Fusobacterium_ species _nucleatum_ and _varium_ have been commonly isolated from mucosal biopsies of patients with IBD @Lee2016. Laboratory experiments with these isolates have shown that disease-isolated _F. nucleatum_ are more invasive and stimulate more TNF-$\alpha$ production than strains from healthy individuals @Strauss2011, suggesting the bacteria may increase inflammation in the gut as well @Dharmani2011. _F. varium_ isolated from UC patients caused colonic ulcers in an experimental mouse model @Ohkusa2003. _F. varium_ was only detected in three subjects and two of those samples were isolated form the proximal mucosa (Supplementary Fig. S1B).  _F. varium_ is most commonly isolated from UC patient biopsies from the ileum or cecum @Ohkusa2002, suggesting this species may exhibit preference for the different environmental conditions of these gastrointestinal sites. Further work will assess how gut environment may select for species which may then cause localized disease.

>>>>

We detected _F. nucleatum_ and _P. asacharolytica_ in 8 and 5 of our subjects, respectively. These bacteria have been shown to be predictive of colorectal cancer in humans @Baxter2016 and have oncogenic properties in cell culture and in mice @Sears2014. Though the bacteria are known to co-localize on CRC tumors, in our study _F. nucleatum_ was found on both sides of the colon while _P. asacharolytica_ was only detected in the distal mucosa. Not much is known about the distribution of _P. asacharolytica_ along the healthy colon, but given its anaerobic lifestyle and asacharolytic metabolism, it is perhaps not surprising that our study detected the bacteria primarily in the less-oygen-rich and protein-rich distal mucosa @Albenberg2014. In studies examining bacteria on colorectal cancer tumors, _F. nucleatum_ was more commonly detected on proximal-sided tumors, and distribution of _F. nucleatum_ decreased along the colon to rectum @Mima2016. Of the 8 (40%) individuals positive for _F. nucleatum_ in our study, the bacterium was spread across the proximal mucosa, distal lumen and distal mucosa. The _Fusobacterium_ species _nucleatum_ and _varium_ have been commonly isolated from mucosal biopsies of patients with IBD and UC [@Lee2016; @Ohkusa2003]. In our study, _F. varium_ was only detected in three subjects and two of those samples were isolated form the proximal mucosa (Supplementary Fig. S1B).  _F. varium_ is most commonly isolated from UC patient biopsies from the ileum or cecum (adjacent to the proximal colon) @Ohkusa2002, suggesting this species may exhibit preference for the higher oxygen content of these gastrointestinal sites.

The distribution of these disease-associated microbes in healthy patients is of interest their presence can be predictive of disease prior to the onset of symptoms @Baxter2016. Spatial organization of _Fusobacteria_ and other bacterial species into polymicrobial biofilms that can invade the gut mucosa have been linked CRC (cite new paper). The organization into biofilms promotes tumorigenesis by allowing bacteria to grow near the epithelium, promoting inflammation, genotoxicity and metabolic changes that favor tumor cell growth (cite new paper and kaitlin). In one study of CRC biofilms and tumors, all of the proximal tumors examined contained a polymicrobial biofilm on the tumor mucosa @Dejea2014. Further examination of these tumors identified _Fusobacteria_ species as members of the biofilm. These results indicate that it is not only the presence of the bacterium but the tumor community as a whole that contributes to tumorigenesis @Dejea2014. Driver passeneger model (cite new paper and KF).  Need to study healthy people/biofilms more to understand how can affect disease. Further analysis of the distribution and activities of these pathogens along the colon is needed to elucidate a mechanism for development of CRC or IBD subtypes in the proximal or distal colon.

paragraph on bacterial organization/spatial variation along colon vs at tumor site/biofilms
